\ieeetitle

### 2017-02-06 report/abstract.tex
Correct typos in report

\begin{abstract}
  \epto  \autocite{matos2015epto} is one of the recently introduced total order algorithms for large-scale distributed systems, and provides total order and scalability at the same time. In this paper, we verify this claim by designing a framework to facilitate the evaluation of large scale distributed systems, implement \epto and evaluate its reliability in real-world conditions by comparing it to a deterministic total order algorithm named \jgroups. We first compare them in a perfect environment assessing the scalability in terms of events throughput and peer numbers. We then compare them using a synthetic churn trace and a real one using the same metrics. Our results show \epto is performing has expected, although further testing on a bigger cluster is required to evaluate how it performs when \jgroups no longer does.
  \hm{General comment: if possible, maybe we should rent resources on Amazon to test with larger networks.}
\end{abstract}

>>>>

\begin{abstract}
  \epto  \autocite{matos2015epto} is one of the recently introduced total order algorithms for large-scale distributed systems, and provides total order and scalability at the same time. In this paper, we verify this claim by designing a framework to facilitate the evaluation of large scale distributed systems, implement \epto and evaluate its reliability in real-world conditions by comparing it to a deterministic total order algorithm named \jgroups. We first compare them in a perfect environment assessing the scalability in terms of events throughput and peer numbers. We then compare them using a synthetic churn trace and a real one using the same metrics. Our results show \epto is performing as expected, although further testing on a bigger cluster is required to evaluate how it performs when \jgroups no longer does.
  \hm{General comment: if possible, maybe we should rent resources on Amazon to test with larger networks.}
\end{abstract}

### 2017-02-06 report/background.tex
Correct typos in report

\section{Background}
\label{sec:background}
Distributed systems and centralized systems alike, need to preserve the temporal order of events produced by concurrent processes in the system. When there are separated processes that can only communicate through messages, you cannot easily order these messages.
Therefore we need ordering algorithms to overcome this problem.
\par
We have two types of ordering algorithms \autocite{lamport1978time}: the partial order algorithms and the total order algorithms.
\subsection{Partial Order Algorithms}
Assuming S is partially ordered under $\leq$, then the following statements hold for all a, b and c in S:
\begin{itemize}
	\item Reflexivity: $a \leq a$ for all $a \in S$.
	\item Antisymmetry: $a \leq b$ and $b \leq a$ implies $a=b$ .
	\item Transitivity: $a \leq b$  and $b \leq c$  implies $a \leq c$.
\end{itemize}

@@

Another protocol named \textit{PABCast} \autocite{felber2002} provides almost the same properties as \epto. However \textit{PABCast} only offers probabilistic order whereas \epto offers deterministic order. Furthermore, the standard \textit{PABCast} protocol only allows for peers to broadcast one message per round whereas \epto allows for an indefinite number of messages. This restriction can be overcome with \textit{PABCast} extensions, but the probability of having different orders among peers increases.

Both \autocite{Chandra2007} and \autocite{Maia2011} extensively test a particular large scale distributed system. However, they do not offer a framework to benchmark different systems. \autocite{Leonini2009} presents a framework named SPlay to deploy such applications. Unfortunately, SPlay only allows for LUA applications to run. \eptotester offers a convenient framework to deploy distributed applications regardless of the programming language used, provided they run in a docker container. This makes it easier for existing protocol implementations to be benchmarked using our framework.

>>>>

\section{Background}
\label{sec:background}
Distributed systems and centralized systems alike need to preserve the temporal order of events produced by concurrent processes in the system. When there are separated processes that can only communicate through messages, you cannot easily order these messages.
Therefore we need ordering algorithms to overcome this problem.
\par
We have two types of ordering algorithms \autocite{lamport1978time}: the partial order algorithms and the total order algorithms.
\subsection{Partial Order Algorithms}
Assuming S is partially ordered under $\leq$, then the following statements hold for all a, b and c in S:
\begin{itemize}
	\item Reflexivity: $a \leq a$ for all $a \in S$.
	\item Antisymmetry: $a \leq b$ and $b \leq a$ implies $a=b$ .
	\item Transitivity: $a \leq b$  and $b \leq c$  implies $a \leq c$.
\end{itemize}

@@

Another protocol named \textit{PABCast} \autocite{felber2002} provides almost the same properties as \epto. However \textit{PABCast} only offers probabilistic order whereas \epto offers deterministic order. Furthermore, the standard \textit{PABCast} protocol only allows for peers to broadcast one message per round whereas \epto allows for an indefinite number of messages. This restriction can be overcome with \textit{PABCast} extensions, but the probability of having different orders among peers increases, making it less desirable than \epto.

Both \autocite{Chandra2007} and \autocite{Maia2011} extensively test a particular large scale distributed system. However, they do not offer a framework to benchmark different systems or use SPlay \autocite{Leonini2009}. SPlay is a framework used to deploy distributed applications. Unfortunately, SPlay only allows for LUA applications to run. \eptotester offers a convenient framework to deploy distributed softare regardless of the programming language used, provided they run in a docker container. This makes it easier for existing protocol implementations to be benchmarked using our framework.

### 2017-02-06 report/conclusion.tex
Correct typos in report

In this paper, we implement \eptotester an architecture designed to facilitate the deployment of a large-scale distributed system on a cluster and the injection of churn in the network. Furthermore, we implement \epto and evaluate the claims made in \autocite{matos2015epto} against \jgroups SEQUENCER, a total order algorithm relying on a coordinator. Unfortunately, our cluster cannot handle more than 100 nodes at the same time. Due to this limitation, \epto performs worse than \jgroups. We want to reiterate that this is expected as \epto is designed for large distributed networks. \jt{Let's write it this way for the moment if we are able to run something on GCE or AWS we can always change it later.}

>>>>

In this paper, we implement \eptotester a framework designed to facilitate the deployment of a large-scale distributed system on a cluster and the injection of churn in the network. Furthermore, we implement \epto and evaluate the claims made in \autocite{matos2015epto} against \jgroups SEQUENCER, a total order algorithm relying on a coordinator. Unfortunately, our cluster cannot handle more than 100 nodes at the same time. Due to this limitation, \epto performs worse than \jgroups. We want to reiterate that this is expected as \epto is designed for large distributed networks.

### 2017-02-06 report/evaluation.tex
Correct typos in report

	\item[\textbf{$(100,50)$}:] 100 peers with a global event throughput of 50 events per second. These parameters were also used for all synthetic churns.

@@

\textbf{Testbed.} \mm{this sentence is not necessary} \jt{I specified it as the following text is really close to the one in Sebastien work as we use the same cluster and I don't want any problem being accused of copy}Our work uses the same cluster first used in \autocite{vaucher2016erasure}. It consists of 20 machines interconnected by a \SI{1}{\gbps} switched network. Each machine has an 8-core Intel Xeon CPU and \SI{8}{\giga\byte} of RAM. We deploy 12 virtual machines on these hosts. Each virtual machine has access 4 VCPUs as well as \SI{7}{\giga\byte} of RAM. We use KVM as our hypervisor. Each VM uses Debian as its O/S. \eptotester is packaged as docker images. We use Docker 1.12 and its new functionality Docker Swarm to orchestrate our services. Each docker container has a memory restriction of \SI{300}{\mega\byte} of RAM to be able to run them all together.

@@

\textbf{Experiment parameters.} Every benchmark except the one following a real trace are run 10 times, each during \SI{20}{\minute}. When there is synthetic churn, the churn starts \SI{30}{\second} after the benchmark and run for \SI{17}{\minute}. The benchmarks following a real trace are run 5 times. The trace is speed-up 2x, which means we follow \SI{2}{\hour} worth of trace in \SI{1}{\hour}. Every benchmark run with churn is run with the $(100,50)$ parameters. When we have a global event throughput lower than the number of actual peers, we use a uniform random number generator to artificially reduce the global event throughput. For example, using $(100,50)$ parameters, each peer has a 50\% probability of sending an event at each second.

@@

Comparing \epto and \jgroups in terms of bandwidth when we increase the number of peers, \epto scales better than \jgroups. Where \jgroups basically has to double the bandwidth usage of the coordinator, \epto only increases it marginally \jt{logarithmic I think as was expected}. Thus in a scenario where we have many peers \epto will be more efficient than \jgroups at not reaching the maximum available bandwidth on a given node.\mm{it would be perfect if we could experimentally show this.}
 \begin{figure}[hpt]
 	\centering
 	\includegraphics[width=\linewidth]{figures/bandwidth-synth-churn.pdf}
 	\vspace{-2mm} 
 	\caption{Bandwidth percentiles of nodes during an unstable experiment with synthetic churn}
 	\vspace{-2mm} 
 	\label{fig:bandwidth-churn}
 \end{figure}
\par
In \autoref{fig:bandwidth-churn} We analyze two different synthetic churns. In the first one we kill one node per minute. In the second one, we still kill one node per minute, but we immediately create a new one. For \jgroups we run the benchmarks once without killing the coordinator and once killing it.

We can see that the churn does not affect \epto at all when there are only nodes leaving. We have small peaks when adding a node to \SI{3}{\mbps} or less. Probably due to running the PSS initialization method on top of having one more node spreading rumors in the system. This is confirmed at the end of the plot where \epto goes back to a normal Bandwidth after stabilization.

@@

\subsection{Total GigaBytes sent/received}
This metric shows the Total number of bytes sent and received on average during an experiment.
In all cases, \jgroups receives more bytes than it sends. This is due to the fact that we do not measure the bandwidth on the TCPGOSSIP, which is in charge of view changes. Thus, when a new peer joins the cluster, the TCPGOSSIP will send a new view to every peer already in the cluster.
\par
\begin{table}[hpt]
	\centering
	\caption{Total \si{\giga\byte} sent/received in a stable system}
	\sisetup{table-format=2.2, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
	\resizebox{\linewidth}{!}{
	\begin{tabular}{llSSS}
		\toprule
		&& \multicolumn{3}{c}{Cluster parameters} \\
		\cmidrule{3-5}
		{Protocol}&& {$(50,50)$} & {$(50,100)$} & {$(100,50)$} \\
		\midrule
		\multirow{2}{*}{\epto}&{Receive}& 10.84(016) & 22.31(039) & 26.01(027) \\
						    &{Sending}& 10.84(016) & 22.31(039) & 26.01(027)\\
		\midrule
		\multirow{2}{*}{\jgroups}&{Receive}& 0.78(003) & 1.45(001) & 1.88(001)\\
							   &{Sending}& 0.77(003) & 1.44(001) & 1.84(001)\\
		\bottomrule
	\end{tabular}
	}
	\label{table:total-bandwidth} 
\end{table}
In \autoref{table:total-bandwidth}, \epto has a worse baseline than \jgroups. This is expected as \epto sends $c*n*\log_2 n$ messages per events and \jgroups sends at least $n$ messages per event so we should have at least $c*\log_2 n$ more messages sent in \epto if \jgroups is perfect. Here we are well within this ratio.
\begin{table}[hpt]
	\centering
	\caption{Total \si{\giga\byte} sent/received with a synthetic churn}
	\sisetup{table-format=2.2, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
	\resizebox{\linewidth}{!}{
	\begin{tabular}{lSSS}
		\toprule
		&& \multicolumn{2}{c}{Churn parameters} \\
		\cmidrule{3-4}
		{Protocol}&& {1 kill/minute} & {1\{kill,add\}/minute} \\
		\midrule
		\multirow{2}{*}{\epto}&{Receive}& 21.00(024) & 26.32(032)\\
							&{Sending}& 21.21(025) & 26.57(032)\\
		\midrule
		\multirow{2}{*}{\jgroups-coord}&{Receive}& 1.47(002) & 1.75(002)\\&{Sending}& 1.43(002) & 1.70(002)\\
		\midrule
		\multirow{2}{*}{\jgroups-nocoord}&{Receive}& 1.45(001) & 1.73(002)\\&{Sending}& 1.41(001) & 1.68(002)\\
		\bottomrule
	\end{tabular}
	}
	\label{table:total-bandwidth-churn} 
\end{table}
In \autoref{table:total-bandwidth-churn} we see that \jgroups total bandwidth usage is smaller when there is churn. One hypothesis for this is that a \jgroups replica takes a longer time to start up compared to stopping a replica. Therefore the overall benchmark has a longer time with less than 100 replicas. We also do not see a difference whether we kill the coordinator or not. This can be explained by the fact that before the detection of a faulty coordinator \jgroups is forced to a halt for period of up to \SI{20}{\second}. The big spike afterwards compensates for this hole.
\subsection{Local Times}
\label{sub:local-times}
The local time represents the total time it takes for a peer to deliver every events sent.
 \begin{figure*}[hpt]
 	\centering
 	\includegraphics[width=\linewidth]{figures/local-times-nochurn.pdf}
 	\vspace{-2mm} 
 	\caption{Local dissemination times}
 	\vspace{-2mm}
 	\label{fig:local-times} 
 \end{figure*}

 \begin{figure}[hpt]
 	\centering
 	\includegraphics[width=\linewidth]{figures/local-times-synth-churn.pdf}
 	\vspace{-2mm} 
 	\caption{Local dissemination times with churn}
 	\vspace{-2mm} 
 	\label{fig:local-times-churn} 
 \end{figure}
In \autoref{fig:local-times}, \jgroups delivers all events quicker than \epto in all scenarios, even when churn is involved as is shown in \autoref{fig:local-times-churn}. However, \epto is not too far behind. The difference between \epto and \jgroups is likely to be even smaller when running them in a real WAN network due to the latency. \epto in our configuration has a $\delta$ period of \SI{100}{\milli\second} and is thus handicapped against \jgroups in a LAN environment, because it only increments the TTL of an event every \SI{100}{\milli\second}. We expect \epto to outperform \jgroups in this situation when we have a high number of peers in the system or when the network has a higher latency.\mm{we could try to setup an AWS test in mutliple availability zones to see what happens.} \jt{Yes I though about it too. Unfortunately we can't make it work for the moment :/}
\subsection{Global Times}
The global time represents the total time it takes for all peers to deliver every events sent.
 \begin{figure*}[hpt]
 	\centering
 	\includegraphics[width=\linewidth]{figures/global-times-nochurn.pdf}
 	\vspace{-2mm} 
 	\caption{Global dissemination times}
 	\vspace{-2mm}
 	\label{fig:global-times}  
 \end{figure*}

@@

On the contrary, \epto performs very well under churn. The median degrades a bit at \SI{650}{\milli\second} with the 99th percentile being at \SI{1030}{\milli\second} compared to \SI{982}{\milli\second} when no churn is happening.
\subsection{Events sent}
The variation observed in the different tests are due to the fact that we run an experiment for a period of time and that in some configurations a randomness decides whether an event is sent or not.
\begin{table}[hpt]
	\centering
	\caption{Total events sent in a stable environment}
\sisetup{table-format=6.1, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
\begin{tabular}{lSSS}
	\toprule
	& \multicolumn{3}{c}{Cluster parameters} \\
	\cmidrule{2-4}
	Protocol & {$(50,50)$} & {$(50,100)$} & {$(100,50)$} \\
	\midrule
	\epto & 59993.8(33) & 119898.2(97) & 59913.0(1643) \\
	\jgroups & 59961.9(109) & 119885.7(50) & 60023.1(2871) \\
	\bottomrule
\end{tabular}
\label{table:total-events}  
\end{table}

@@

In \autoref{fig:local-delta-real-churn}, \epto as worse outliers than before and so does \jgroups-coord.

@@

	Protocol &\\

@@

When running \jgroups we sometimes get failed runs, where either some peers will experience holes or peer will deliver out of order events. We measure  These failed runs only happen when there is churn happening, specifically when we add and remove peers and when the coordinator dies. These problems appeared 3 times out of 13 runs for \jgroups $1\{kill,add\}/minute$ and 2 times out of 7 for \jgroups with real churn. We have not investigated these problems further. This points to either \jgroups failing when there is too much churn or a bug present in the SEQUENCER implementation.

We measure holes as one or more events that were delivered by at least one peer in the cluster but not on the peer containing the hole. As stated earlier, we specifically choose a $c$ high enough to never experience holes in \epto as \jgroups should never suffer from holes. If either a \jgroups or \epto fails to deliver every event sent during the run, the run is considered as having a problem.

>>>>

	\item[\textbf{$(100,50)$}:] 100 peers with a global event throughput of 50 events per second. These parameters are also used for all synthetic churns.

@@

\textbf{Testbed.} \mm{this sentence is not necessary} \jt{I specified it as the following text is really close to the one in Sebastien work as we use the same cluster and I don't want any problem being accused of copy}Our work uses the same cluster first used in \autocite{vaucher2016erasure}. It consists of 20 machines interconnected by a \SI{1}{\gbps} switched network. Each machine has an 8-core Intel Xeon CPU and \SI{8}{\giga\byte} of RAM. We deploy 12 virtual machines on these hosts. Each virtual machine has access 4 VCPUs as well as \SI{7}{\giga\byte} of RAM. We use KVM as our hypervisor. Each VM uses Debian as its O/S. Protocols are packaged as docker images. We use Docker 1.12 and its new functionality Docker Swarm to orchestrate our services. Each docker container has a memory restriction of \SI{300}{\mega\byte} of RAM to be certain we will not have out of memory problems.

@@

\textbf{Experiment parameters.} Every benchmark except the ones following a real trace are run 10 times, each during \SI{20}{\minute}. When there is synthetic churn, the churn starts \SI{30}{\second} after the benchmark and run for \SI{17}{\minute}. The benchmarks following a real trace are run 5 times. The trace is sped-up 2 times, which means we follow \SI{2}{\hour} worth of trace in \SI{1}{\hour}. Every benchmark run with churn is run with the $(100,50)$ parameters. When we have a global event throughput lower than the number of actual peers, we use a uniform random number generator to artificially reduce the global event throughput. For example, using $(100,50)$ parameters, each peer has a 50\% probability of sending an event at each second.

@@

Comparing \epto and \jgroups in terms of bandwidth when we increase the number of peers, \epto scales better than \jgroups. While \jgroups basically has to double the bandwidth usage of the coordinator, \epto only increases it marginally. Thus in a scenario where we have many peers \epto will be more efficient than \jgroups at not reaching the maximum available bandwidth on a given node.\mm{it would be perfect if we could experimentally show this.}
 \begin{figure}[hpt]
 	\centering
 	\includegraphics[width=\linewidth]{figures/bandwidth-synth-churn.pdf}
 	\vspace{-2mm} 
 	\caption{Bandwidth percentiles of nodes during an unstable experiment with synthetic churn}
 	\vspace{-2mm} 
 	\label{fig:bandwidth-churn}
 \end{figure}
\par
In \autoref{fig:bandwidth-churn} We analyze two different synthetic churns. In the first one we kill one node per minute. In the second one, we still kill one node per minute, but we immediately create a new one. For \jgroups we run the benchmarks once without killing the coordinator and once killing it.

We can see that the churn does not affect \epto at all when there are only nodes leaving. We have small peaks when adding a node to \SI{3}{\mbps} or less. Probably due to running the PSS initialization method. This is confirmed at the end of the plot where \epto goes back to a normal Bandwidth after stabilization.

@@

\subsection{Total GigaBytes sent/received}
This metric shows the Total number of bytes sent and received on average during an experiment.
In all cases, \jgroups receives more bytes than it sends. This is due to the fact that we do not measure the bandwidth on the TCPGOSSIP, which is in charge of view changes. Thus, when a new peer joins the cluster, the TCPGOSSIP will send a new view to every peer already in the cluster.
\par
\begin{table}[hpt]
	\centering
	\caption{Total \si{\giga\byte} sent/received in a stable system}
	\sisetup{table-format=2.2, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
	\resizebox{\linewidth}{!}{
	\begin{tabular}{llSSS}
		\toprule
		&& \multicolumn{3}{c}{Cluster parameters} \\
		\cmidrule{3-5}
		{Protocol}&& {$(50,50)$} & {$(50,100)$} & {$(100,50)$} \\
		\midrule
		\multirow{2}{*}{\epto}&{Receive}& 10.84(016) & 22.31(039) & 26.01(027) \\
						    &{Sending}& 10.84(016) & 22.31(039) & 26.01(027)\\
		\midrule
		\multirow{2}{*}{\jgroups}&{Receive}& 0.78(003) & 1.45(001) & 1.88(001)\\
							   &{Sending}& 0.77(003) & 1.44(001) & 1.84(001)\\
		\bottomrule
	\end{tabular}
	}
	\label{table:total-bandwidth} 
\end{table}
In \autoref{table:total-bandwidth}, \epto has a worse baseline than \jgroups. This is expected as \epto sends $c*n*\log_2 n$ messages per events and \jgroups sends at least $n$ messages per event so we should have at least $c*\log_2 n$ more messages sent in \epto if \jgroups is perfect. Here we are well within this ratio.
\begin{table}[hpt]
	\centering
	\caption{Total \si{\giga\byte} sent/received with a synthetic churn}
	\sisetup{table-format=2.2, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
	\resizebox{\linewidth}{!}{
	\begin{tabular}{lSSS}
		\toprule
		&& \multicolumn{2}{c}{Churn parameters} \\
		\cmidrule{3-4}
		{Protocol}&& {1 kill/minute} & {1\{kill,add\}/minute} \\
		\midrule
		\multirow{2}{*}{\epto}&{Receive}& 21.00(024) & 26.32(032)\\
							&{Sending}& 21.21(025) & 26.57(032)\\
		\midrule
		\multirow{2}{*}{\jgroups-coord}&{Receive}& 1.47(002) & 1.75(002)\\&{Sending}& 1.43(002) & 1.70(002)\\
		\midrule
		\multirow{2}{*}{\jgroups-nocoord}&{Receive}& 1.45(001) & 1.73(002)\\&{Sending}& 1.41(001) & 1.68(002)\\
		\bottomrule
	\end{tabular}
	}
	\label{table:total-bandwidth-churn} 
\end{table}
In \autoref{table:total-bandwidth-churn} we see that \jgroups total bandwidth usage is smaller when there is churn. One hypothesis for this is that a \jgroups replica takes a longer time to start up compared to stopping a replica. Therefore the overall benchmark has a longer time with less than 100 replicas. We also do not see a difference whether we kill the coordinator or not. This can be explained by the fact that before the detection of a faulty coordinator \jgroups is forced to a halt for period of up to \SI{20}{\second}. The big spike afterwards compensates for this hole.
\subsection{Local Times}
\label{sub:local-times}
The local time represents the total time it takes for a peer to deliver every .
 \begin{figure*}[hpt]
 	\centering
 	\includegraphics[width=\linewidth]{figures/local-times-nochurn.pdf}
 	\vspace{-2mm} 
 	\caption{Local dissemination times}
 	\vspace{-2mm}
 	\label{fig:local-times} 
 \end{figure*}

 \begin{figure}[hpt]
 	\centering
 	\includegraphics[width=\linewidth]{figures/local-times-synth-churn.pdf}
 	\vspace{-2mm} 
 	\caption{Local dissemination times with churn}
 	\vspace{-2mm} 
 	\label{fig:local-times-churn} 
 \end{figure}
In \autoref{fig:local-times}, \jgroups delivers all events quicker than \epto in all scenarios, even when churn is involved as is shown in \autoref{fig:local-times-churn}. However, \epto is not too far behind. The difference between \epto and \jgroups is likely to be even smaller when running them in a real WAN network due to the latency. \epto in our configuration has a $\delta$ period of \SI{100}{\milli\second} and is thus handicapped against \jgroups in a LAN environment, because it only increments the TTL of an event every \SI{100}{\milli\second}. We expect \epto to outperform \jgroups in this situation when we have a high number of peers in the system or when the network has a higher latency. Furthermore, in our test environment \jgroups is not limited by the bandwidth. If the coordinator would have a lower limit of bandwidth, the time to send all events when the coordinator dies would be longer.\mm{we could try to setup an AWS test in mutliple availability zones to see what happens.} \jt{Yes I though about it too. Unfortunately we can't make it work for the moment :/}
\subsection{Global Times}
The global time represents the total time it takes for all peers to deliver every events sent.
 \begin{figure*}[hpt]
 	\centering
 	\includegraphics[width=\linewidth]{figures/global-times-nochurn.pdf}
 	\vspace{-2mm} 
 	\caption{Global dissemination times}
 	\vspace{-2mm}
 	\label{fig:global-times}  
 \end{figure*}

@@

On the contrary, \epto performs very well under churn. The median degrades a bit at \SI{650}{\milli\second} with the 99th percentile being at \SI{1030}{\milli\second} compared to \SI{982}{\milli\second} when no churn is happening.
\subsection{Events sent}
The variation observed in the different tests are due to the fact that we use \textit{Thread.sleep()} between each send and that in some configurations a randomness decides whether an event is sent or not.
\begin{table}[hpt]
	\centering
	\caption{Total events sent in a stable environment}
\sisetup{table-format=6.1, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
\begin{tabular}{lSSS}
	\toprule
	& \multicolumn{3}{c}{Cluster parameters} \\
	\cmidrule{2-4}
	Protocol & {$(50,50)$} & {$(50,100)$} & {$(100,50)$} \\
	\midrule
	\epto & 59993.8(33) & 119898.2(97) & 59913.0(1643) \\
	\jgroups & 59961.9(109) & 119885.7(50) & 60023.1(2871) \\
	\bottomrule
\end{tabular}
\label{table:total-events}  
\end{table}

@@

In \autoref{fig:local-delta-real-churn}, \epto has worse outliers than before and so does \jgroups-coord.

@@

	Protocol & {Events sent}\\

@@

When running \jgroups we sometimes get failed runs, where either some peers will experience holes or a peer will deliver out of order events. These failed runs only happen when there is churn happening, specifically when we add and remove peers and when the coordinator dies. These problems appeared 3 times out of 13 runs for \jgroups $1\{kill,add\}/minute$ and 2 times out of 7 for \jgroups with real churn. We have not investigated these problems further. This points to either \jgroups failing when there is too much churn or a bug present in the SEQUENCER implementation.

We measure holes as one or more events that were delivered by at least one peer in the cluster but not on the peer containing the hole. As stated earlier, we specifically choose a $c$ high enough to never experience holes in \epto as \jgroups should never suffer from holes. If either \jgroups or \epto fails to deliver every event sent during the run, the run is considered as having a problem.

### 2017-02-06 report/figures/no-churn/bandwidth-figure.tex
WIP report
Correct typos in report

legend style={at={(1.9,1.5)},anchor=north, font=\small, draw=none},]

@@

\node[anchor=south, rotate=-90] at (group c3r1.east){EpTO};
\node[anchor=south, rotate=-90] at (group c3r2.east){JGroups};

>>>>

legend style={at={(1.8,1.5)},anchor=north, font=\small, draw=none, fill=none},]

@@

\node[anchor=south, rotate=-90] at (group c3r1.east){\epto};
\node[anchor=south, rotate=-90] at (group c3r2.east){\jgroups};

### 2017-02-06 report/figures/no-churn/global-times.tex
WIP report
Correct typos in report

\begin{tikzpicture}
\begin{groupplot}[
group style={
	group size=3 by 1,
	vertical sep=0pt,
	horizontal sep=6mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	xticklabels at=edge bottom,
	yticklabels at=edge left,
},
every axis plot/.append style={very thick},
width=\linewidth / 3,
height=5cm,
grid=major,
grid style={dashed,gray!30},
ymax=1.05,
ymin=0,
xmin=1185,
xmax=1202.5,
xlabel={Time $\left[\si{\second}\right]$},
x tick label style={/pgf/number format/fixed},
ytick={0,0.2,0.4,0.6,0.8,1},
yticklabels={0,0.2,0.4,0.6,0.8,1},
ylabel=Percentiles,
legend columns=3,
legend cell align=left,
legend style={at={(1.6,1.3)},anchor=north, font=\small, draw=none},
cycle list name=my colors,
]
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-50-1sec-x, y=EpTO-50-1sec-y, col sep=comma] {\tableglobaltime};
\addplot+[const plot, mark=none] table[x=JGroups-50-1sec-x, y=JGroups-50-1sec-y, col sep=comma] {\tableglobaltime};
\legend{EpTO, JGroups}
%
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-50-2sec-x, y=EpTO-50-2sec-y, col sep=comma] {\tableglobaltime};
\addplot+[const plot, mark=none] table[x=JGroups-50-2sec-x, y=JGroups-50-2sec-y, col sep=comma] {\tableglobaltime};
%
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-100-1sec-x, y=EpTO-100-1sec-y, col sep=comma] {\tableglobaltime};
\addplot+[const plot, mark=none] table[x=JGroups-100-1sec-x, y=JGroups-100-1sec-y, col sep=comma] {\tableglobaltime};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {$(50,50)$};
\node[anchor=south] at (group c2r1.north) {$(50,100)$};
\node[anchor=south] at (group c3r1.north) {$(100,50)$};
\end{tikzpicture}

>>>>

\begin{tikzpicture}
\begin{groupplot}[
group style={
	group size=3 by 1,
	vertical sep=0pt,
	horizontal sep=6mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	xticklabels at=edge bottom,
	yticklabels at=edge left,
},
every axis plot/.append style={very thick},
width=\linewidth / 3,
height=5cm,
grid=major,
grid style={dashed,gray!30},
ymax=1.05,
ymin=0,
xmin=1185,
xmax=1202.5,
xlabel={Time $\left[\si{\second}\right]$},
x tick label style={/pgf/number format/fixed},
ytick={0,0.2,0.4,0.6,0.8,1},
yticklabels={0,0.2,0.4,0.6,0.8,1},
ylabel=Percentiles,
legend columns=3,
legend cell align=left,
legend style={at={(1.6,1.3)},anchor=north, font=\small, draw=none, fill=none},
cycle list name=my colors,
]
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-50-1sec-x, y=EpTO-50-1sec-y, col sep=comma] {\tableglobaltime};
\addplot+[const plot, mark=none] table[x=JGroups-50-1sec-x, y=JGroups-50-1sec-y, col sep=comma] {\tableglobaltime};
\legend{\epto, \jgroups}
%
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-50-2sec-x, y=EpTO-50-2sec-y, col sep=comma] {\tableglobaltime};
\addplot+[const plot, mark=none] table[x=JGroups-50-2sec-x, y=JGroups-50-2sec-y, col sep=comma] {\tableglobaltime};
%
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-100-1sec-x, y=EpTO-100-1sec-y, col sep=comma] {\tableglobaltime};
\addplot+[const plot, mark=none] table[x=JGroups-100-1sec-x, y=JGroups-100-1sec-y, col sep=comma] {\tableglobaltime};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {$(50,50)$};
\node[anchor=south] at (group c2r1.north) {$(50,100)$};
\node[anchor=south] at (group c3r1.north) {$(100,50)$};
\end{tikzpicture}

### 2017-02-06 report/figures/no-churn/local-delta.tex
Correct typos in report

\begin{tikzpicture}
\begin{groupplot}[
group style={
	group size=3 by 1,
	vertical sep=0pt,
	horizontal sep=4mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	xticklabels at=edge bottom,
	yticklabels at=edge left,
},
every axis plot/.append style={very thick},
width=\linewidth / 2.5,
height=5cm,
grid=major,
grid style={dashed,gray!30},
ymax=1.05,
ymin=0,
xmax=3.5,
xlabel={Time $\left[\si{\second}\right]$},
x tick label style={/pgf/number format/fixed},
ytick={0,0.2,0.4,0.6,0.8,1},
yticklabels={0,0.2,0.4,0.6,0.8,1},
ylabel=Percentiles,
legend columns=3,
legend cell align=left,
legend style={at={(1.6,1.3)},anchor=north, font=\small, draw=none},
cycle list name=my colors,
]
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-50-1sec-x, y=EpTO-50-1sec-y, col sep=comma] {\tablelocaldeltas};
\addplot+[const plot, mark=none] table[x=JGroups-50-1sec-x, y=JGroups-50-1sec-y, col sep=comma] {\tablelocaldeltas};
\legend{\epto, \jgroups}
%
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-50-2sec-x, y=EpTO-50-2sec-y, col sep=comma] {\tablelocaldeltas};
\addplot+[const plot, mark=none] table[x=JGroups-50-2sec-x, y=JGroups-50-2sec-y, col sep=comma] {\tablelocaldeltas};
%
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-100-1sec-x, y=EpTO-100-1sec-y, col sep=comma] {\tablelocaldeltas};
\addplot+[const plot, mark=none] table[x=JGroups-100-1sec-x, y=JGroups-100-1sec-y, col sep=comma] {\tablelocaldeltas};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {$(50,50)$};
\node[anchor=south] at (group c2r1.north) {$(50,100)$};
\node[anchor=south] at (group c3r1.north) {$(100,50)$};
\end{tikzpicture}

>>>>

\begin{tikzpicture}
\begin{groupplot}[
group style={
	group size=3 by 1,
	vertical sep=0pt,
	horizontal sep=4mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	xticklabels at=edge bottom,
	yticklabels at=edge left,
},
every axis plot/.append style={very thick},
width=\linewidth / 2.5,
height=5cm,
grid=major,
grid style={dashed,gray!30},
ymax=1.05,
ymin=0,
xmax=3.5,
xlabel={Time $\left[\si{\second}\right]$},
x tick label style={/pgf/number format/fixed},
ytick={0,0.2,0.4,0.6,0.8,1},
yticklabels={0,0.2,0.4,0.6,0.8,1},
ylabel=Percentiles,
legend columns=3,
legend cell align=left,
legend style={at={(1.6,1.3)},anchor=north, font=\small, draw=none, fill=none},
cycle list name=my colors,
]
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-50-1sec-x, y=EpTO-50-1sec-y, col sep=comma] {\tablelocaldeltas};
\addplot+[const plot, mark=none] table[x=JGroups-50-1sec-x, y=JGroups-50-1sec-y, col sep=comma] {\tablelocaldeltas};
\legend{\epto, \jgroups}
%
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-50-2sec-x, y=EpTO-50-2sec-y, col sep=comma] {\tablelocaldeltas};
\addplot+[const plot, mark=none] table[x=JGroups-50-2sec-x, y=JGroups-50-2sec-y, col sep=comma] {\tablelocaldeltas};
%
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-100-1sec-x, y=EpTO-100-1sec-y, col sep=comma] {\tablelocaldeltas};
\addplot+[const plot, mark=none] table[x=JGroups-100-1sec-x, y=JGroups-100-1sec-y, col sep=comma] {\tablelocaldeltas};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {$(50,50)$};
\node[anchor=south] at (group c2r1.north) {$(50,100)$};
\node[anchor=south] at (group c3r1.north) {$(100,50)$};
\end{tikzpicture}
